---
title: "simulation"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# Part 1: Simulation Exercise Instructions
Simulating exponential distribution variables and comparing them with the Central Limit Theorem.

```{r Simulation, echo=FALSE}
lambda<-0.2
TheoMean<-1/lambda
TheoSD<-1/lambda
n<-40
B<-10000
simData<-NULL
for (i in 1:B) {
    simData<-rbind(simData,c(rexp(n, lambda)))
    
}

simMean<-apply(simData, 1, mean)
```

The theoretical mean and SD of an exponential variable is 1/lambda

```{r lambda}
lambda;
1/lambda;
```

to compare to the CLT generate n exponential variables then find the mean of the n results. 
the same action will be executed B times. then a mean of all the means will be calculated and compered to the theoretical value.

this are the parameters we are going to use: 
```{r parameters}
n;
B;
```
The simulated mean is: 
```{r simMean, echo=FALSE}
mean(simMean);
```
The difference between the theoretical mean and the simulated is: 
```{r mean_diff, echo=FALSE}
TheoMean - mean(simMean);
```
The simulated SD is: 
```{r SD, echo=FALSE}
sd(simMean);
```
and the theoretical SD is:
```{r lambda2}
1/lambda;
```

The distrebution of the simulated mean is approximetly normal with mean 1/lambda as possible to see in the histogrem. the red line  denotes the theoretical mean. 1/lambda.

```{r hist, echo=FALSE }
hist(simMean,main = "Histogram of the exponential means");
abline(v=TheoMean, col = 'Red',lwd = 2)
lines(density(X)) 
```
